---
title: "Taxdata-DataBC Integration"
author: "Noushin Nabavi"
date: '2019-05-09'
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Taxdata integration with Data Catalogue

- Load dependencies for the project: 

```{r source, results="hide", include=FALSE, echo = FALSE}
## Source setup and function scripts
if (!exists(".setup_sourced")) source(here::here("R/setup.R"))
```


- Load the data: 

```{r data.table, eval = TRUE}
## load tax data table
ind_1 <- fread(here("input-data", "1_IND.csv"))
```



### exploring levels of geography:

Levels of geography in taxdata include:    
-  3 FSA: ([URBAN FSA](https://www12.statcan.gc.ca/census-recensement/2016/ref/dict/geo036-eng.cfm))   
-  6 RURPC: (RURAL POSTAL CODE AREAS)   
-  7 OUA: (OTHER URBAN AREAS)   
- 8 CITY: (CITY TOTAL)   
- 9 RURC: (RURAL COMMUNITIES)   
- 10 OPROV: (OTHER PROVINCIAL TOTAL)   
- 11 PROV: ([PROVINCE/TERRITORY TOTAL](https://www12.statcan.gc.ca/census-recensement/2016/ref/dict/geo038-eng.cfm))   
- 12 CANADA: ([CANADA](https://www12.statcan.gc.ca/census-recensement/2016/ref/98-304/chap12-eng.cfm))   
- 21 CD: ([CENSUS DIVISION](https://www12.statcan.gc.ca/census-recensement/2016/ref/dict/geo008-eng.cfm))   
- 31 FED: ([FEDERAL ELECTORAL DISTRICT](https://www12.statcan.gc.ca/census-recensement/2016/ref/dict/geo025-eng.cfm))   
- 51 ER: ([ECONOMIC REGION](https://www12.statcan.gc.ca/census-recensement/2016/ref/dict/geo022-eng.cfm))   
- 41 CMA: ([CENSUS METROPOLITAN AREA](https://www12.statcan.gc.ca/census-recensement/2016/ref/dict/geo009-eng.cfm))   
- 42 CA: ([CENSUS AGGLOMERATION](https://www12.statcan.gc.ca/census-recensement/2016/ref/dict/geo005-eng.cfm))   
- 61 CT: ([CENSUS TRACT](https://www12.statcan.gc.ca/census-recensement/2016/ref/dict/geo013-eng.cfm))   

### Data Catalogue queries for geography levels:

```{r census, eval = FALSE}
# explore data catalogue using keywords
bcdata::bcdc_browse()
bcdata::bcdc_search("census")

```

- load relevant data from the catalogue: 
```{r bcdc, echo = TRUE}

# Current Census Subdivision Boundaries
CSD <- bcdata::bcdc_get_data("4c5618c6-38dd-4a62-a3de-9408b4974bb6")
mapview(CSD)

# Census Tracts
CT <- bcdata::bcdc_get_data("539aae5b-12f6-4934-9592-9b27acc827f8")
mapview(CT)

# Current Census Metropolitan Areas (other, wms, kml)
CMA <- bcdata::bcdc_get_data("a6fb34b7-0937-4718-8f1f-43dba2c0f407")
mapview(CMA)

# Current Census Division Boundaries (other, wms, kml)
CD <- bcdata::bcdc_get_data("ef17918a-597a-4012-8534-f8e71d8735b3")
mapview(CD)

# Current Census Economic Regions (other, wms, kml)
ER <- bcdata::bcdc_get_data("1aebc451-a41c-496f-8b18-6f414cde93b7")
mapview(ER)

```
