---
title: "Taxdata-DataBC Integration"
author: "Noushin Nabavi"
date: '2019-05-09'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Taxdata integration with Data Catalogue

### Levels of geography:

Levels of geography in taxdata include:    
-  3 FSA: ([URBAN FSA](https://www12.statcan.gc.ca/census-recensement/2016/ref/dict/geo036-eng.cfm))   
-  6 RURPC: (RURAL POSTAL CODE AREAS)   
-  7 OUA: (OTHER URBAN AREAS)   
- 8 CITY: (CITY TOTAL)   
- 9 RURC: (RURAL COMMUNITIES)   
- 10 OPROV: (OTHER PROVINCIAL TOTAL)   
- 11 PROV: ([PROVINCE/TERRITORY TOTAL](https://www12.statcan.gc.ca/census-recensement/2016/ref/dict/geo038-eng.cfm))   
- 12 CANADA: ([CANADA](https://www12.statcan.gc.ca/census-recensement/2016/ref/98-304/chap12-eng.cfm))   
- 21 CD: ([CENSUS DIVISION](https://www12.statcan.gc.ca/census-recensement/2016/ref/dict/geo008-eng.cfm))   
- 31 FED: ([FEDERAL ELECTORAL DISTRICT](https://www12.statcan.gc.ca/census-recensement/2016/ref/dict/geo025-eng.cfm))   
- 51 ER: ([ECONOMIC REGION](https://www12.statcan.gc.ca/census-recensement/2016/ref/dict/geo022-eng.cfm))   
- 41 CMA: ([CENSUS METROPOLITAN AREA](https://www12.statcan.gc.ca/census-recensement/2016/ref/dict/geo009-eng.cfm))   
- 42 CA: ([CENSUS AGGLOMERATION](https://www12.statcan.gc.ca/census-recensement/2016/ref/dict/geo005-eng.cfm))   
- 61 CT: ([CENSUS TRACT](https://www12.statcan.gc.ca/census-recensement/2016/ref/dict/geo013-eng.cfm))   

### Data Catalogue queries for geography levels:

```{r census, eval = FALSE}
bcdata::bcdc_browse()

bcdata::bcdc_search("census")

census_sub <- bcdata::bcdc_get_data("4c5618c6-38dd-4a62-a3de-9408b4974bb6")
max.plot(census_sub)
```

