---
title: "Taxdata-DataBC Integration"
date: '2019-05-09'
output:
  html_document: default
  pdf_document: default
---

```{r setup, echo = FALSE, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# About tax data 
The technical Reference Guide for the Annual Income Estimates for Census Families, Individuals and Seniors can be found in the `/docs` folder. For information about this product or the wide range of services and data available from Statistics Canada, visit their [website](www.statcan.gc.ca).
For further questions, the Statistics Canada email is <STATCAN.infostats-infostats.STATCAN@canada.ca>. 


- Load dependencies for the project by loading `setup.R` script: 

```{r source, results="hide", include=FALSE, echo = FALSE}
## Source setup and function scripts
if (!exists(".setup_sourced")) source(here::here("R/setup.R"))
```

# Levels of geography

Levels of geography included in the tax data are:  

-  3 FSA: ([URBAN FSA](https://www12.statcan.gc.ca/census-recensement/2016/ref/dict/geo036-eng.cfm))   
-  6 RURPC: (RURAL POSTAL CODE AREAS)   
-  7 OUA: (OTHER URBAN AREAS)   
- 8 CITY: (CITY TOTAL)   
- 9 RURC: (RURAL COMMUNITIES)   
- 10 OPROV: (OTHER PROVINCIAL TOTAL)   
- 11 PROV: ([PROVINCE/TERRITORY TOTAL](https://www12.statcan.gc.ca/census-recensement/2016/ref/dict/geo038-eng.cfm))   
- 12 CANADA: ([CANADA](https://www12.statcan.gc.ca/census-recensement/2016/ref/98-304/chap12-eng.cfm))   
- 21 CD: ([CENSUS DIVISION](https://www12.statcan.gc.ca/census-recensement/2016/ref/dict/geo008-eng.cfm))   
- 31 FED: ([FEDERAL ELECTORAL DISTRICT](https://www12.statcan.gc.ca/census-recensement/2016/ref/dict/geo025-eng.cfm))   
- 51 ER: ([ECONOMIC REGION](https://www12.statcan.gc.ca/census-recensement/2016/ref/dict/geo022-eng.cfm))   
- 41 CMA: ([CENSUS METROPOLITAN AREA](https://www12.statcan.gc.ca/census-recensement/2016/ref/dict/geo009-eng.cfm))   
- 42 CA: ([CENSUS AGGLOMERATION](https://www12.statcan.gc.ca/census-recensement/2016/ref/dict/geo005-eng.cfm))   
- 61 CT: ([CENSUS TRACT](https://www12.statcan.gc.ca/census-recensement/2016/ref/dict/geo013-eng.cfm))   


Based on initial consultations and research using Statistics Canada as source, the levels of geography that are usable for this project include those that are small enough to hold the smallest number of taxfilers in a region and is inclusive of all of BC. The challenge here is that not one geography level can be chosen such that it covers all of BC. We determined that economic indices need to be generated for rural and urban communities separately. In the case that a postal code does not fall in the urban (census tracts) or rural (rural communities and postal code) geographies, then we take the next high level of geography (i.e. census division). More specifically, geo level #9 is for RURAL COMMUNITIES, geo level #6 is for RURAL POSTAL CODE AREAS, and geo level #21 is for CENSUS DIVISION.

# Workflow for economic index generation

The simplest strategy for quintile ratio generation is determined by dividing the median total income (from table 1 of individuals) into 5 quintile ranges on a per year and region basis. The quintile ratio varies every year and depending on the regions of geography and the median total income of taxfilers in that region. 

In order to generate the indices, we synthesized a fake data that resembles administrative datasets. Every studyid is associated with a random postal code and level of geography with primary keys. The synthesized data is then linked to tax data in order to associate every region to a quintile range per year. The codes for synthesizing data and making economic indices are in `R` subfolder, named `synthesize_data.R` and `indicies.R` respectively. The urban and rural indices designated for the synthesized data are named as urban quintiles (UQ) and rural quintiles (RQ) and are placed in the `output-data` folder.


```{r income, echo= FALSE, eval=TRUE}
library(knitr)
library(kableExtra)
Market_Income <- data.frame(
  # = 1:4,	
  
  Market_Income_Structure = c(
"Employment income",
"Investment income",
"Private retirement income",
"Market income not included elsewhere"), 

Categories = c(
"Wages, salaries, commissions, self-employment",
"-",
"-",
"-"),
 stringsAsFactors = FALSE)
``` 

```{r income image, echo=FALSE, eval=TRUE}
kable(cbind(Market_Income), "latex", booktabs = T, caption = "Market Income Structures") %>%
  kable_styling(latex_options = c("striped", "scale_down", font_size=200))
```


```{r gov income, echo =FALSE, eval=TRUE}
library(knitr)
library(kableExtra)
Government_Income <- data.frame(
  # = 1:5,	
  
Government_Transfer_Structure = c(
"Old Age Security pension (OAS)",
"Canada and Québec Pension Plan (CPP-QPP) benefits",
"Employment Insurance (EI) benefits",
"Child benefits",
"Other government transfers"),

Categories = c(
"-",
"-",
"EI-retirement, disability, survivor benefits", 
"Canada, Provincial and Territorial child tax benefits",
"Social assistance, Workers' compensation, GST-HST tax credits, others"
),																																																				
 stringsAsFactors = FALSE)
``` 

```{r income gov image, echo=FALSE, eval=TRUE}
kable(cbind(Government_Income), "latex", booktabs = T, caption = "Government Income Structures") %>%
  kable_styling(latex_options = c("striped", "scale_down", font_size=200))
```


```{r  individual tables, echo= FALSE, eval=TRUE}
library(knitr)
library(kableExtra)
library(dplyr)
Individuals <- data.frame(
  # = 1:19,	
  
  Table_ID = c("Table I-01", "Table I-02", "Table I-3A", "Table I-3B", "Table I-3C", "Table I-04", "Table I-5A", "Table I-5B",
            "Table I-5C", "Table I-06", "Table I-7A", "Table I-7B", "Table I-7C", "Table I-08", "Table I-09", "Table I-10",
            "Table I-11", "Table I-12", "Table I-13"), 
  
  Table_Name = c("Taxfilers and dependents, summary table, income and demographics of individuals", "Taxfilers and dependents by sex, marital status and age groups", "Male taxfilers and dependents by single years of age", "Female taxfilers and dependents by single years of age", "Taxfilers and dependents by single year of age", "Taxfilers and dependents with income by source of income", "Male taxfilers and dependents with income by total Income and age groups", "Female taxfilers and dependents with income by total income and age groups", "Taxfilers and dependents with income by total income and age groups", "Taxfilers and dependents with income by sex, income taxes, selected deductions and benefits", "Male taxfilers and dependents with income by after-tax income and age groups", "Female taxfilers and dependents  with income by after-tax income and age groups", "Taxfilers and dependents with income by after-tax income and age groups", "Taxfilers and dependents with income by income taxes and after-tax income, sex and age groups", "Economic dependency profile of individuals", "Labour income profile of individuals", "Taxfilers and dependents 15 years of age and over with labour income by sex and age groups", "Taxfilers and dependents 15 years of age and over receiving employment insurance by age groups and sex", "After-tax low income status of tax filers and dependents, census family low income measure, for couple and lone parent families by family composition"),																																																																											
 stringsAsFactors = FALSE)
``` 

```{r individual table image, echo= FALSE, eval=TRUE}
kable(cbind(Individuals), "latex", booktabs = T, caption = "Structure for Individual Tables") %>%
  kable_styling(latex_options = c("striped", "scale_down", font_size=200))
```


```{r family tables, echo= FALSE, eval=TRUE}
library(knitr)
library(dplyr)
library(kableExtra)
Family <- data.frame(
  # = 1:26,	
  
  Table_ID = c("Table F-01", "Table F-02", "Table F-3A", "Table F-3B", "Table F-3C", "Table F-04A", "Table F-04B", "Table F-04C", "Table F-5A", "Table F-5B", "Table F-06", "Table F-7", "Table F-08", "Table F-09", "Table F-10", "Table F-11", "Table F-12", "Table F-13", "Table F-14A", "Table F-14B", "Table F-14C", "Table F-15", "Table F-17", "Table F-18", "Table F-19", "Table F-20"),
  
  Table_Name = c("Summary census family income table", "Taxfilers and dependents by age groups and census family type", "Couple families by age of older partner or parent and number of children", "Lone-Parent families by age of parent and number of children", "Census families by age of older partner or parent and number of children", "Distribution of total income by couple family and age of older partner or parent", "Distribution of total income by lone-parent family and age of parent", "Distribution of total income of persons not in census families by age of individual", "Couple families by total income and number of children", "Lone-Parent families by total income and number of children", "Sources of Income by census family Type", "Economic dependency profile of couple families", "Economic dependency profile of lone-parent families and persons not in census families", "Labour income profile of couple families", "Labour income profile of lone-parent families and persons not in census families", "Labour income by age group and sex", "Employment insurance by age group and sex", "Single-earner and dual-earner families by number of children, includes only partners, parents reporting non-negative employment income", "Couple families by percentage of wife's contribution to couple's employment income and by number of children", "Couple families by percentage of wife's contribution to husband/wife employment income and by age of wife", "Couple families by percentage of wife's contribution to husband, wife employment income and by family employment income range", "Census families with children by age of children and children by age groups", "Before-tax low income status, based on census family low income measures, family type and family composition", "After-tax low income status, based on census family low income measures, by family type and family composition", "After-tax low income status of census families, census family low income measures, by family type and family composition, adjusted methodology", "Census families by family type and family composition including before and after-tax median income of the family"),
  
  
 stringsAsFactors = FALSE)
``` 


```{r tamily table image, echo= FALSE, eval=TRUE}
kable(cbind(Family), "latex", booktabs = T, caption = "Structure for Family Tables") %>%
  kable_styling(latex_options = c("striped", "scale_down", font_size=200))
```

# Resources	
- Dictionary, Census of Population, 2016
- Guide to the Census of Population, 2016
- Income Reference Guide, Census of Population, 2016
- Statistics Canada's Open Licence
- Reference Guide for the Annual Income Estimates for Census Families, Individuals and Seniors 2016
- A Review of Methods for Deriving an Index for Socioeconomic Status in British Columbia April 2013
- Postal Code OM Conversion File (PCCF), Reference Guide, June 2017

# Appendix I: 
Income inequality in Canada is defined in more detail, [here](https://www.compareyourincome.org/income_inequality_in_canada#what_is_the_issue). 
Based on this, the average income in Canada stands at 44000 Canadian dollars. The total income explorer released by Statistics Canada provides additional measures, [here](https://www12.statcan.gc.ca/census-recensement/2016/dp-pd/dv-vd/inc-rev/index-eng.cfm), and [here](https://www150.statcan.gc.ca/n1/daily-quotidien/180322/dq180322b-eng.htm). 
Additional socioeconomic reports can be found, [here](https://www12.statcan.gc.ca/nhs-enm/2011/as-sa/99-014-x/99-014-x2011003_2-eng.cfm).


