---
title: "Taxdata-DataBC Integration"
date: '2019-05-09'
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Taxdata integration with Data Catalogue
Income Estimates for Census Families, Individuals and Seniors can be found under STATSCAN_METADATA_2016. For information about this data or the wide range of services and data available from Statistics Canada, visit their [website](www.statcan.gc.ca). You can also contact Statistics Canada by emailing STATCAN.infostats-infostats.STATCAN@canada.ca.

- Load dependencies for the project: 

```{r source, results="hide", include=FALSE, echo = FALSE}
## Source setup and function scripts
if (!exists(".setup_sourced")) source(here::here("R/setup.R"))
```

# Workflow for economic index generation:
The workflow for generating an economic index consists of devising metrics based on available StatCan annual taxdata at individual and family levels. The following document contains fields from individual tables and their potential usage for index generation. 


# Income structures in the data


```{r income, echo= TRUE, eval=TRUE}
library(knitr)
library(kableExtra)
Market_Income <- data.frame(
  # = 1:4,	
  
  Market_Income_Structure = c(
"Employment income",
"Investment income",
"Private retirement income",
"Market income not included elsewhere"), 

Categories = c(
"Wages, salaries, commissions, self-employment",
"-",
"-",
"-"),
 stringsAsFactors = FALSE)
``` 

```{r income image, echo= TRUE, eval=TRUE}
kable(cbind(Market_Income), "latex", booktabs = T, caption = "Market Income Structures") %>%
  kable_styling(latex_options = c("striped", "scale_down", font_size=200))
```


```{r gov income, echo= TRUE, eval=TRUE}
library(knitr)
library(kableExtra)
Government_Income <- data.frame(
  # = 1:5,	
  
Government_Transfer_Structure = c(
"Old Age Security pension (OAS)",
"Canada and Québec Pension Plan (CPP-QPP) benefits",
"Employment Insurance (EI) benefits",
"Child benefits",
"Other government transfers"),

Categories = c(
"-",
"-",
"EI-retirement, disability, survivor benefits", 
"Canada, Provincial and Territorial child tax benefits",
"Social assistance, Workers' compensation, GST-HST tax credits, others"
),																																																				
 stringsAsFactors = FALSE)
``` 

```{r income gov image, echo= TRUE, eval=TRUE}
kable(cbind(Government_Income), "latex", booktabs = T, caption = "Government Income Structures") %>%
  kable_styling(latex_options = c("striped", "scale_down", font_size=200))
```


```{r  individual tables, echo= TRUE, eval=TRUE}
library(knitr)
library(kableExtra)
Individuals <- data.frame(
  # = 1:19,	
  
  Table_ID = c("Table I-01", "Table I-02", "Table I-3A", "Table I-3B", "Table I-3C", "Table I-04", "Table I-5A", "Table I-5B",
            "Table I-5C", "Table I-06", "Table I-7A", "Table I-7B", "Table I-7C", "Table I-08", "Table I-09", "Table I-10",
            "Table I-11", "Table I-12", "Table I-13"), 
  
  Table_Name = c("Taxfilers and dependents, summary table, income and demographics of individuals", "Taxfilers and dependents by sex, marital status and age groups", "Male taxfilers and dependents by single years of age", "Female taxfilers and dependents by single years of age", "Taxfilers and dependents by single year of age", "Taxfilers and dependents with income by source of income", "Male taxfilers and dependents with income by total Income and age groups", "Female taxfilers and dependents with income by total income and age groups", "Taxfilers and dependents with income by total income and age groups", "Taxfilers and dependents with income by sex, income taxes, selected deductions and benefits", "Male taxfilers and dependents with income by after-tax income and age groups", "Female taxfilers and dependents  with income by after-tax income and age groups", "Taxfilers and dependents with income by after-tax income and age groups", "Taxfilers and dependents with income by income taxes and after-tax income, sex and age groups", "Economic dependency profile of individuals", "Labour income profile of individuals", "Taxfilers and dependents 15 years of age and over with labour income by sex and age groups", "Taxfilers and dependents 15 years of age and over receiving employment insurance by age groups and sex", "After-tax low income status of tax filers and dependents, census family low income measure, for couple and lone parent families by family composition"),																																																																											
 stringsAsFactors = FALSE)
``` 

```{r individual table image, echo= TRUE, eval=TRUE}
kable(cbind(Individuals), "latex", booktabs = T, caption = "Structure for Individual Tables") %>%
  kable_styling(latex_options = c("striped", "scale_down", font_size=200))
```


```{r family tables, echo= TRUE, eval=TRUE}
library(knitr)
library(kableExtra)
Family <- data.frame(
  # = 1:26,	
  
  Table_ID = c("Table F-01", "Table F-02", "Table F-3A", "Table F-3B", "Table F-3C", "Table F-04A", "Table F-04B", "Table F-04C", "Table F-5A", "Table F-5B", "Table F-06", "Table F-7", "Table F-08", "Table F-09", "Table F-10", "Table F-11", "Table F-12", "Table F-13", "Table F-14A", "Table F-14B", "Table F-14C", "Table F-15", "Table F-17", "Table F-18", "Table F-19", "Table F-20"),
  
  Table_Name = c("Summary census family income table", "Taxfilers and dependents by age groups and census family type", "Couple families by age of older partner or parent and number of children", "Lone-Parent families by age of parent and number of children", "Census families by age of older partner or parent and number of children", "Distribution of total income by couple family and age of older partner or parent", "Distribution of total income by lone-parent family and age of parent", "Distribution of total income of persons not in census families by age of individual", "Couple families by total income and number of children", "Lone-Parent families by total income and number of children", "Sources of Income by census family Type", "Economic dependency profile of couple families", "Economic dependency profile of lone-parent families and persons not in census families", "Labour income profile of couple families", "Labour income profile of lone-parent families and persons not in census families", "Labour income by age group and sex", "Employment insurance by age group and sex", "Single-earner and dual-earner families by number of children, includes only partners, parents reporting non-negative employment income", "Couple families by percentage of wife's contribution to couple's employment income and by number of children", "Couple families by percentage of wife's contribution to husband/wife employment income and by age of wife", "Couple families by percentage of wife's contribution to husband, wife employment income and by family employment income range", "Census families with children by age of children and children by age groups", "Before-tax low income status, based on census family low income measures, family type and family composition", "After-tax low income status, based on census family low income measures, by family type and family composition", "After-tax low income status of census families, census family low income measures, by family type and family composition, adjusted methodology", "Census families by family type and family composition including before and after-tax median income of the family"),
  
  
 stringsAsFactors = FALSE)
``` 


```{r tamily table image, echo= TRUE, eval=TRUE}
kable(cbind(Family), "latex", booktabs = T, caption = "Structure for Family Tables") %>%
  kable_styling(latex_options = c("striped", "scale_down", font_size=200))
```

# Resources	
- Dictionary, Census of Population, 2016
- Guide to the Census of Population, 2016
- Income Reference Guide, Census of Population, 2016
- Statistics Canada's Open Licence
- Reference Guide for the Annual Income Estimates for Census Families, Individuals and Seniors 2016
- A Review of Methods for Deriving an Index for Socioeconomic Status in British Columbia April 2013
- Postal Code OM Conversion File (PCCF), Reference Guide, June 2017


### exploring levels of geography:

Levels of geography in taxdata include:  

-  3 FSA: ([URBAN FSA](https://www12.statcan.gc.ca/census-recensement/2016/ref/dict/geo036-eng.cfm))   
-  6 RURPC: (RURAL POSTAL CODE AREAS)   
-  7 OUA: (OTHER URBAN AREAS)   
- 8 CITY: (CITY TOTAL)   
- 9 RURC: (RURAL COMMUNITIES)   
- 10 OPROV: (OTHER PROVINCIAL TOTAL)   
- 11 PROV: ([PROVINCE/TERRITORY TOTAL](https://www12.statcan.gc.ca/census-recensement/2016/ref/dict/geo038-eng.cfm))   
- 12 CANADA: ([CANADA](https://www12.statcan.gc.ca/census-recensement/2016/ref/98-304/chap12-eng.cfm))   
- 21 CD: ([CENSUS DIVISION](https://www12.statcan.gc.ca/census-recensement/2016/ref/dict/geo008-eng.cfm))   
- 31 FED: ([FEDERAL ELECTORAL DISTRICT](https://www12.statcan.gc.ca/census-recensement/2016/ref/dict/geo025-eng.cfm))   
- 51 ER: ([ECONOMIC REGION](https://www12.statcan.gc.ca/census-recensement/2016/ref/dict/geo022-eng.cfm))   
- 41 CMA: ([CENSUS METROPOLITAN AREA](https://www12.statcan.gc.ca/census-recensement/2016/ref/dict/geo009-eng.cfm))   
- 42 CA: ([CENSUS AGGLOMERATION](https://www12.statcan.gc.ca/census-recensement/2016/ref/dict/geo005-eng.cfm))   
- 61 CT: ([CENSUS TRACT](https://www12.statcan.gc.ca/census-recensement/2016/ref/dict/geo013-eng.cfm))   


Based on initial scans of data and maps, the levels of geography that are usable for this project include those that are small enough to hold the smallest number of taxfilers in a region and is inclusive of all of BC. The challenge here is that not one geography level can be chosen and so  economic indices needs to be generated for rural and urban communities separately. 

Another challenge is that the same city, e.g. Abbotsford, can repeat multiple times with the same level of geography but different postal regions. 
Since a cityID is also provided, we need to roll up the taxfilers information based on both level of geography and common cityIDs.

1. Levels of geography:

Geography levels need to be used for index generation since not all regions are covered by one geo level alone. Geo level #9 is for RURAL COMMUNITIES and #61 is for CENSUS TRACT. RURAL POSTAL CODE AREAS (#6) is also usable as they are small enough and exclusive of rural communities. CENSUS DIVISION (#21) also provides small enough regions <50,000 which is exclusive of census tracts.

2. cityID: 

Since levels of geography and postal names can have repetitions, we would need to consider names of places, geography and postal codes, as well as cityID to roll-up the taxfilers' information into one row per region of interest. Stat Can and PCCF metadata also refer to usage of cityIDs in decoding postal areas.

3. quintile ratio:

The simplest strategy for quintile ratio generation is dividing the income for urban and rural regions into 5 ratios. Currenly, the income is divided in <5k+, 5k+, 10k+, 15k+, 20k+, 25k+, 35k+, 50k+, 75k+, 75k+, 100k+, 150k+, 200k+, 250k+. 

A quintile ratio, however, can vary depending on the regions of geography and the normal distribution of taxfilers in each region. 

The index will vary depending on the distribution of taxfilers based on year and region.



# Data usage:

The simplest linkage will include use of individual table #1 from taxdata. This table holds information about taxfilers and dependents, summary table, income and demographics of individuals. Individual table #5 is the next simplest table for more detailed index generation that sheds light on gender and age distribution as well as income. Tables #6-12 include additional complexities and details that can be used depending on ministry needs.


```{r script, results="hide", include=FALSE, echo = FALSE}

## Source setup and function scripts
if (!exists(".setup_sourced")) source(here::here("R/setup.R"))
```


# Appendix I: 
Individual Taxdata Tables structures:

### Table I-01
This table holds information about taxfilers and dependents, summary table, income and demographics of individuals.
 
```{r table 1, eval=FALSE}
ind_1 <- fread(here("input-data", "1_IND.csv"))

Fields relevant to index:
  
postal|area|
year
cityid|
level|of|geo|
place|me|geo|
taxfilers|#|
total|income|median|males
total|income|median|females
total|income|median|total

```


### Table I-02
This table includes information on taxfilers and dependents by sex, marital status and age groups. 
There are **no** fields for index generation, directly.

```{r table 2, eval=FALSE}
ind_2 <- fread(here("input-data", "2_IND.csv"))
```

### Table I-3A-B-C
These tables includes information on male (3A) and female (3B) and combined (3C) taxfilers and dependents by single years of age. 
There are **no** fields for index generation, directly.

```{r table 3, eval=FALSE}
ind_3A <- fread(here("input-data", "3A_IND.csv"))
ind_3B <- fread(here("input-data", "3B_IND.csv"))
ind_3C <- fread(here("input-data", "3C_IND.csv"))
```


### Table I-04 
This table includes information on taxfilers and dependents with income by source of income.
There are **no** fields for index generation, directly.

```{r table 4, eval=FALSE}
ind_4 <- fread(here("input-data", "4_IND.csv"))
```

### Table I-5A-B-C
These tables include male, female, and all taxfilers and dependents with income by total Income and age groups.
The breakdowns for age and income can help with index generation.

```{r table 5, eval=FALSE}
ind_5A <- fread(here("input-data", "5A_IND.csv"))
ind_5B <- fread(here("input-data", "5B_IND.csv"))
ind_5C <- fread(here("input-data", "5C_IND.csv"))

Fields of interest include:
  
postal|area|	
cityid|	
level|of|geo|	
place|me|
  ## taxfilers in the brackets below:
  all|males|with|income|by|total|income|0-24|years|<$5k
	all|males|with|income|by|total|income|0-24|years|$5k+	   
	all|males|with|income|by|total|income|0-24|years|$10k+
	all|males|with|income|by|total|income|0-24|years|$15k+
	all|males|with|income|by|total|income|0-24|years|$20k+
	all|males|with|income|by|total|income|0-24|years|$25k+
	all|males|with|income|by|total|income|0-24|years|$35k+
	all|males|with|income|by|total|income|0-24|years|$50k+
	all|males|with|income|by|total|income|0-24|years|$75k+
	all|males|with|income|by|total|income|0-24|years|$100k+
	all|males|with|income|by|total|income|0-24|years|$150k+
	all|males|with|income|by|total|income|0-24|years|$200k+
	all|males|with|income|by|total|income|0-24|years|$250k+
	all|males|with|income|by|total|income|0-24|years|total

- Other age brackets include:
  25-34, 35-44, 45-54, 55-64, 65-74, 75+

- This information also exist for females and both genders combined.
```


### Table I-06 
Taxfilers and dependents with income by sex, income taxes, selected deductions and benefits.

```{r table 6, eval=FALSE}
ind_6 <- fread(here("input-data", "6_IND.csv"))

postal|area|
year	
cityid|	
level|of|geo|	
place|me|
  
total|income|$'000|males
total|income|$'000|females
total|income|$'000|total
```

### Table I-7A-B-C
This table includes information on male, female, and all taxfilers and dependents with income by after-tax income and age groups

```{r table 7, eval=FALSE}
ind_7 <- fread(here("input-data", "7_IND.csv"))

postal|area|	
year	
cityid|	
level|of|geo|	
place|me|
    ## taxfilers in the brackets below:
  all|males|with|after-tax|income|by|total|income|0-24|years|<$5k
  all|males|with|after-tax|income|by|total|income|0-24|years|$5k+	
  all|males|with|after-tax|income|by|total|income|0-24|years|$10k+
  all|males|with|after-tax|income|by|total|income|0-24|years|$15k+	
  all|males|with|after-tax|income|by|total|income|0-24|years|$20k+  
  all|males|with|after-tax|income|by|total|income|0-24|years|$25k+
  all|males|with|after-tax|income|by|total|income|0-24|years|$35k+
  all|males|with|after-tax|income|by|total|income|0-24|years|$50k+
  all|males|with|after-tax|income|by|total|income|0-24|years|$75k+
  all|males|with|after-tax|income|by|total|income|0-24|years|$100k+   
  all|males|with|after-tax|income|by|total|income|0-24|years|$150k+
  all|males|with|after-tax|income|by|total|income|0-24|years|$200k+
  all|males|with|after-tax|income|by|total|income|0-24|years|$250k+
  all|males|with|after-tax|income|by|total|income|0-24|years|total
  
Other age brackets include:
  25-34, 35-44, 45-54, 55-64, 65-74, 75+

- This information also exist for females and both genders combined.

```


### Table I-08 
This table includes information on taxfilers and dependents with income by income taxes and after-tax income, sex and age groups.

```{r table 8, eval=FALSE}
ind_8 <- fread(here("input-data", "8_IND.csv"))

postal|area|
year	
cityid|		
postal|walk|	
level|of|geo|	
place|me|
  ## taxfilers in the brackets below:
  all|males|total|income|#|0-24|years	
  all|males|total|income|#|$'000|0-24|years

- Other age brackets include:
  25-34, 35-44, 45-54, 55-64, 65+

- This information also exist for females and both genders combined.
```

### Table I-09 
This table includes economic dependency profile of individuals.

```{r table 9, eval=FALSE}
ind_9 <- fread(here("input-data", "9_IND.csv"))

postal|area|	
year	
cityid|
level|of|geo|
place|me|
empl.|income|#|males	
empl.|income|#|females	
empl.|income|#|total	
empl.|income|$'000|males	
empl.|income|$'000|females	
empl.|income|$'000|total

- Other sources of income include: 
- government and non-government transfers.
```

### Table I-10 
Labour income profile of individuals in this table can be used for index generation:

```{r table 10, eval=FALSE}
ind_10 <- fread(here("input-data", "10_IND.csv"))
postal|area|
year	
cityid|	
level|of|geo|	
place|me|
all|persons|males|	
all|persons|females|	
all|persons|total|	
taxfilers|males|	
taxfilers|females|	
taxfilers|total|	
total|income|number|males	
total|income|number|females	
total|income|number|total	
total|income|$'000|males	
total|income|$'000|females	
total|income|$'000|total	
total|income|median|males	
total|income|median|females	
total|income|median|total	

- Other sources of income include: 
labour, employment, wages, salaries, self employment, income, 
```

### Table I-11 
Table including taxfilers and dependents at 15 years of age and over with labour income by sex and age groups. Fields of interest include:

```{r table 11, eval=FALSE}
ind_11 <- fread(here("input-data", "11_IND.csv"))
postal|area|
year	
cityid|	
level|of|geo|	
place|me|
  
- Other sources of income include: 
- female and male age ranges within income brackets
```

### Table I-12 
Table includes taxfilers and dependents at 15 years of age and over receiving employment insurance by age groups and sex. Fields of interest include:

```{r table 12, eval=FALSE}
ind_12 <- fread(here("input-data", "12_IND.csv"))

postal|area|
year	
cityid|	
level|of|geo|	
place|me|

- Other sources of income include: 
- labour, employment insurance income by age and sex categories 
```

### Table I-13 
Table includes after-tax low income status of tax filers and dependents, census family low income measure, for couple and lone parent families by family composition. There are **no** fields for index generation, directly.

```{r table 13, eval=FALSE}
ind_13 <- fread(here("input-data", "13_IND.csv"))
```

# Appendix II: 
Income inequality in Canada is defined in more detail, [here](https://www.compareyourincome.org/income_inequality_in_canada#what_is_the_issue). 
Based on this, the average income in Canada stands at 44000 Canadian dollars. The total income explorer released by Statistics Canada provides additional measures, [here](https://www12.statcan.gc.ca/census-recensement/2016/dp-pd/dv-vd/inc-rev/index-eng.cfm), and [here](https://www150.statcan.gc.ca/n1/daily-quotidien/180322/dq180322b-eng.htm). 
Additional socioeconomic reports can be found, [here](https://www12.statcan.gc.ca/nhs-enm/2011/as-sa/99-014-x/99-014-x2011003_2-eng.cfm).


# Resources	
- Dictionary, Census of Population, 2016
- Guide to the Census of Population, 2016
- Income Reference Guide, Census of Population, 2016
- Statistics Canada's Open Licence
- Reference Guide for the Annual Income Estimates for Census Families, Individuals and Seniors 2016
- A Review of Methods for Deriving an Index for Socioeconomic Status in British Columbia April 2013
- Postal Code OM Conversion File (PCCF), Reference Guide, June 2017
