---
title: "Taxdata Tables for Index Generation"
date: '2019-05-17'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r script, results="hide", include=FALSE, echo = FALSE}

## Source setup and function scripts
if (!exists(".setup_sourced")) source(here::here("R/setup.R"))
```

## Individual Taxdata Tables:

### Table I-01
This table holds information about Taxfilers and dependents, summary table, income and demographics of individuals.
 
```{r table 1, eval=FALSE}
ind_1 <- fread(here("input-data", "1_IND.csv"))

Fields relevant to index:
  
postal|area|
year
cityid|
level|of|geo|
place|me|geo|
taxfilers|#|
total|income|median|males
total|income|median|females
total|income|median|total

```


### Table I-02
This table includes information on taxfilers and dependents by sex, marital status and age groups. 
There are **no** fields for index generation, directly.

```{r table 2, eval=FALSE}
ind_2 <- fread(here("input-data", "2_IND.csv"))
```

### Table I-3A-B-C
These tables includes information on male (3A) and female (3B) and combined (3C) taxfilers and dependents by single years of age. 
There are **no** fields for index generation, directly.

```{r table 3, eval=FALSE}
ind_3A <- fread(here("input-data", "3A_IND.csv"))
ind_3B <- fread(here("input-data", "3B_IND.csv"))
ind_3C <- fread(here("input-data", "3C_IND.csv"))
```


### Table I-04 
This table includes information on taxfilers and dependents with income by source of income.
There are **no** fields for index generation, directly.

```{r table 4, eval=FALSE}
ind_4 <- fread(here("input-data", "4_IND.csv"))
```

### Table I-5A-B-C
These tables include male, female, and all taxfilers and dependents with income by total Income and age groups.
The breakdowns for age and income can help with index generation.

```{r table 5, eval=FALSE}
ind_5A <- fread(here("input-data", "5A_IND.csv"))
ind_5B <- fread(here("input-data", "5B_IND.csv"))
ind_5C <- fread(here("input-data", "5C_IND.csv"))

Fields of interest include:
  
postal|area|	
cityid|	
level|of|geo|	
place|me|
  ## taxfilers in the brackets below:
  all|males|with|income|by|total|income|0-24|years|<$5k
	all|males|with|income|by|total|income|0-24|years|$5k+	   
	all|males|with|income|by|total|income|0-24|years|$10k+
	all|males|with|income|by|total|income|0-24|years|$15k+
	all|males|with|income|by|total|income|0-24|years|$20k+
	all|males|with|income|by|total|income|0-24|years|$25k+
	all|males|with|income|by|total|income|0-24|years|$35k+
	all|males|with|income|by|total|income|0-24|years|$50k+
	all|males|with|income|by|total|income|0-24|years|$75k+
	all|males|with|income|by|total|income|0-24|years|$100k+
	all|males|with|income|by|total|income|0-24|years|$150k+
	all|males|with|income|by|total|income|0-24|years|$200k+
	all|males|with|income|by|total|income|0-24|years|$250k+
	all|males|with|income|by|total|income|0-24|years|total

Other age brackets include:
  25-34, 35-44, 45-54, 55-64, 65-74, 75+

This information also exist for females and both genders combined.
```


### Table I-06 
Taxfilers and dependents with income by sex, income taxes, selected deductions and benefits.

```{r table 6, eval=FALSE}
ind_6 <- fread(here("input-data", "6_IND.csv"))

postal|area|
year	
cityid|	
level|of|geo|	
place|me|
total|income|$'000|males
total|income|$'000|females
total|income|$'000|total
```

### Table I-7A-B-C
This table includes information on male, female, and all taxfilers and dependents with income by after-tax income and age groups

```{r table 7, eval=FALSE}
ind_7 <- fread(here("input-data", "7_IND.csv"))

postal|area|	
year	
cityid|	
level|of|geo|	
place|me|
    ## taxfilers in the brackets below:
  all|males|with|after-tax|income|by|total|income|0-24|years|<$5k
  all|males|with|after-tax|income|by|total|income|0-24|years|$5k+	
  all|males|with|after-tax|income|by|total|income|0-24|years|$10k+
  all|males|with|after-tax|income|by|total|income|0-24|years|$15k+	
  all|males|with|after-tax|income|by|total|income|0-24|years|$20k+  
  all|males|with|after-tax|income|by|total|income|0-24|years|$25k+
  all|males|with|after-tax|income|by|total|income|0-24|years|$35k+
  all|males|with|after-tax|income|by|total|income|0-24|years|$50k+
  all|males|with|after-tax|income|by|total|income|0-24|years|$75k+
  all|males|with|after-tax|income|by|total|income|0-24|years|$100k+   
  all|males|with|after-tax|income|by|total|income|0-24|years|$150k+
  all|males|with|after-tax|income|by|total|income|0-24|years|$200k+
  all|males|with|after-tax|income|by|total|income|0-24|years|$250k+
  all|males|with|after-tax|income|by|total|income|0-24|years|total
  
Other age brackets include:
  25-34, 35-44, 45-54, 55-64, 65-74, 75+

This information also exist for females and both genders combined.

```


### Table I-08 
This table includes information on taxfilers and dependents with income by income taxes and after-tax income, sex and age groups.

```{r table 8, eval=FALSE}
ind_8 <- fread(here("input-data", "8_IND.csv"))

postal|area|
year	
cityid|		
postal|walk|	
level|of|geo|	
place|me|
  ## taxfilers in the brackets below:
  all|males|total|income|#|0-24|years	
  all|males|total|income|#|$'000|0-24|years

Other age brackets include:
  25-34, 35-44, 45-54, 55-64, 65+

This information also exist for females and both genders combined.
```

### Table I-09 
Economic dependency profile of individuals

```{r table 9, eval=FALSE}
ind_9 <- fread(here("input-data", "9_IND.csv"))
```

### Table I-10 
Labour income profile of individuals

```{r table 10, eval=FALSE}
ind_10 <- fread(here("input-data", "10_IND.csv"))
```

### Table I-11 
Taxfilers and dependents 15 years of age and over with labour income by sex and age groups

```{r table 11, eval=FALSE}
ind_11 <- fread(here("input-data", "11_IND.csv"))
```

### Table I-12 
Taxfilers and dependents 15 years of age and over receiving employment insurance by age groups and sex

```{r table 12, eval=FALSE}
ind_12 <- fread(here("input-data", "12_IND.csv"))
```

### Table I-13 
After-tax low income status of tax filers and dependents, census family low income measure, for couple and lone parent families by family composition

```{r table 13, eval=FALSE}
ind_13 <- fread(here("input-data", "13_IND.csv"))
```


